---
title: 'Assignment 3: Data Exploration'
author: "Yuechen Huang"
date: "Fall 2023"
output:
  pdf_document: default
  html_document:
    df_print: paged
geometry: margin=2.54cm
editor_options:
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Exploration.

## Directions

1.  Rename this file `<FirstLast>_A03_DataExploration.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction. 
4.  Assign a useful **name to each code chunk** and include ample **comments** with your code.
5.  Be sure to **answer the questions** in this assignment document.
6.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
7.  After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai.

**TIP**: If your code extends past the page when knit, tidy your code by manually inserting line breaks.

**TIP**: If your code fails to knit, check that no `install.packages()` or `View()` commands exist in your code. 

---

## Set up your R session

1.  Check your working directory, load necessary packages (tidyverse, lubridate), and upload two datasets: the ECOTOX neonicotinoid dataset (ECOTOX_Neonicotinoids_Insects_raw.csv) and the Niwot Ridge NEON dataset for litter and woody debris (NEON_NIWO_Litter_massdata_2018-08_raw.csv). Name these datasets "Neonics" and "Litter", respectively. Be sure to include the subcommand to read strings in as factors.

```{r set_up_R}
library(tidyverse)
library(lubridate)

Neonics <- read.csv('D:/ENV872_DataExploration/ENV872_DataExploration_Fall2023/Data/Raw/ECOTOX_Neonicotinoids_Insects_raw.csv', stringsAsFactors = TRUE)
Litter <- read.csv('D:/ENV872_DataExploration/ENV872_DataExploration_Fall2023/Data/Raw/NEON_NIWO_Litter_massdata_2018-08_raw.csv', stringsAsFactors = TRUE)
```

## Learn about your system

2.  The neonicotinoid dataset was collected from the Environmental Protection Agency's ECOTOX Knowledgebase, a database for ecotoxicology research. Neonicotinoids are a class of insecticides used widely in agriculture. The dataset that has been pulled includes all studies published on insects. Why might we be interested in the ecotoxicology of neonicotinoids on insects? Feel free to do a brief internet search if you feel you need more background information.

> Answer: the specificity of these neonicotinoids is very important. If these neonicorinoids have a high specificity, other insects (that have important ecological impacts) can still be alive. Human and other mammals will also be safe. The knowing the ecotoxicology helps us to understand the specificity and protect our food and water resources for the current and future generations.


3.  The Niwot Ridge litter and woody debris dataset was collected from the National Ecological Observatory Network, which collectively includes 81 aquatic and terrestrial sites across 20 ecoclimatic domains. 32 of these sites sample forest litter and woody debris, and we will focus on the Niwot Ridge long-term ecological research (LTER) station in Colorado. Why might we be interested in studying litter and woody debris that falls to the ground in forests? Feel free to do a brief internet search if you feel you need more background information.

> Answer: Woody debris has important roles in carbon recycle and can provide habitat to terrestrial and aquatic creatures. Litter debris, however, can be a source of plastic pollution. These litter debris can affect soil quality, negatively influence human and other animals' health. Therefore, studying litter and woody debris is important to forest ecology and health of different creatures.

4.  How is litter and woody debris sampled as part of the NEON network? Read the NEON_Litterfall_UserGuide.pdf document to learn more. List three pieces of salient information about the sampling methods here:

> Answer: 
 1. Litter and fine woody debris sampling is executed at terrestrial NEON sites that contain woody vegetation >2m tall.
 2. Ground traps are sampled once per year.  Target sampling frequency for elevated traps varies by vegetation present at the site, with frequent sampling (1x every 2 weeks) in deciduous forest sites during senescence, and infrequent year-round sampling (1x every 1-2 months) at evergreen sites. 
 3. In sites with forested tower airsheds, the litter sampling is targeted to take place in 20 40m x 40m plots. 

## Obtain basic summaries of your data (Neonics)

5.  What are the dimensions of the dataset?

```{r dimensions_of_dataset}
colnames(Neonics)
# column names refer to the dimensions of dataset
```

6.  Using the `summary` function on the "Effect" column, determine the most common effects that are studied. Why might these effects specifically be of interest?

```{r Effect_of_interest}
summary(Neonics$Effect)
```

> Answer: Population seems to be of interest as 1803 studies are population studies and mortality is the second (1493). The reason why population study is the most popular catagory may be that population study is the basic study for ecotoxicity. Most of the studies need to be done in population level to confirm the toxicity level before further research.


7.  Using the `summary` function, determine the six most commonly studied species in the dataset (common name). What do these species have in common, and why might they be of interest over other insects? Feel free to do a brief internet search for more information if needed.[TIP: The `sort()` command can sort the output of the summary command...]

```{r Neonics_common_name_summary}
summary(Neonics$Species.Common.Name)
sort(summary(Neonics$Species.Common.Name), decreasing = TRUE) # sort the output of species common names in a descending manor
```

> Answer: Honey bees are the most popular research subject because honey bees are ecologically important making them easy to transport pollutants. They are resilient to environmental stress.
References: 
Cunningham MM, Tran L, McKee CG, et al. Honey bees as biomonitors of environmental contaminants, pathogens, and climate change. Ecol Ind. 2022;134:108457. https://www.sciencedirect.com/science/article/pii/S1470160X21011225. doi: 10.1016/j.ecolind.2021.108457.

8.  Concentrations are always a numeric value. What is the class of `Conc.1..Author.` column in the dataset, and why is it not numeric?

```{r concentration_class}
class(Neonics$Conc.1..Author.)
```

> Answer: The class of `Conc.1..Author.` column is factor because some of the values are 'NR' and some values contains '/'.

## Explore your data graphically (Neonics)

9.  Using `geom_freqpoly`, generate a plot of the number of studies conducted by publication year.

```{r freqploy_plot}
plot1_NeoFreq <- ggplot(Neonics) + 
  geom_freqpoly(aes(x = Publication.Year), bins = 15) + theme_light() + labs(x = 'Publication Year', y = 'Number of publication')
plot1_NeoFreq
```

10. Reproduce the same graph but now add a color aesthetic so that different Test.Location are displayed as different colors.

```{r change_color}
plot2_NeoFreqColor <- ggplot(Neonics) + 
  geom_freqpoly(aes(x = Publication.Year, color = Test.Location), bins = 15) + theme_light() + labs(x = 'Publication Year', y = 'Number of publication')
plot2_NeoFreqColor
```

Interpret this graph. What are the most common test locations, and do they differ over time?

> Answer: The most common test locations are lab and field natural. Lab experiements have a trend of increasing from 1980 to around 2014 and decreasing afterwards. Similarly, the number of natual field experiments increases and decreases. The peak of the publications of natural field experiement is around 2009.

11. Create a bar graph of Endpoint counts. What are the two most common end points, and how are they defined? Consult the ECOTOX_CodeAppendix for more information. 

[**TIP**: Add `theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))` to the end of your plot command to rotate and align the X-axis labels...]

```{r bar_graph_of_Endpoint_counts}
plot3_NeoBar <- ggplot(Neonics) + 
  geom_bar(aes(x = Endpoint)) + theme_light() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(y = 'Counts')
plot3_NeoBar
```

> Answer: NOEL and LOEL are the two most common end points. NOEL (No-observable-effect-level) and LOEL (Lowest-observable-effect-level) are common Endpoints for terrestrial database. 

## Explore your data (Litter)

12. Determine the class of collectDate. Is it a date? If not, change to a date and confirm the new class of the variable. Using the `unique` function, determine which dates litter was sampled in August 2018.

```{r litter_collectDate}
class(Litter$collectDate) # the class of collectDate is factor
Litter$collectDate <- as.Date(Litter$collectDate, Format = "%Y-%m-%d") # change factor into date
class(Litter$collectDate) # check the class again
unique(Litter$collectDate) # to see which dates litter was sampled
```

13. Using the `unique` function, determine how many plots were sampled at Niwot Ridge. How is the information obtained from `unique` different from that obtained from `summary`?

```{r plots_sampled_at_Niwot_Ridge}
unique(Litter$fieldSampleID) # to see how many unique values are in the column fieldSampleID
summary(Litter$fieldSampleID)
```

> Answer: Total of 23 plots were sampled at Niwot Ridge. Summary() returns the sample sites but also how many results are in the same site while unique() only returns the name of each unique sample site.

14. Create a bar graph of functionalGroup counts. This shows you what type of litter is collected at the Niwot Ridge sites. Notice that litter types are fairly equally distributed across the Niwot Ridge sites.

```{r litter_bar_graph}
plot4_litterBar <- ggplot(Litter) + 
  geom_bar(aes(x = functionalGroup)) + theme_light() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(y = 'Counts', x = 'Functional Group')
plot4_litterBar
```

15. Using `geom_boxplot` and `geom_violin`, create a boxplot and a violin plot of dryMass by functionalGroup.

```{r litter_dryMass_plot}
# box plot
plot5_litterBox <- ggplot(Litter) + 
  geom_boxplot(aes(x = functionalGroup, y = dryMass)) + theme_light() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(y = 'Dry Mass', x = 'Functional Group')
plot5_litterBox

# violin plot
plot6_litterVio <- ggplot(Litter) + 
  geom_violin(aes(x = functionalGroup, y = dryMass)) + theme_light() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(y = 'Dry Mass', x = 'Functional Group')
plot6_litterVio
```

Why is the boxplot a more effective visualization option than the violin plot in this case?

> Answer: Violin plots may not work well when the sample size is small or there are multiple peaks rather than having a unimodal distribution. Therefore, in our case, the boxplot can be a more effection option which can also clearly show outliers.

What type(s) of litter tend to have the highest biomass at these sites?

> Answer: Needles
